---
title: "CLAV Data Analysis"
author: "Benny deMayo"
date: "June 5, 2015"
output:
  html_document:
    highlight: pygments
    theme: flatly
  pdf_document: default
---

### Intro and Imports
I first clear all previous variables and load libraries for data manipulation and graphing
```{r message=FALSE}
rm(list = ls())

library(tidyverse)
library(lmerTest)
library(fs)


#file paths
project_root <- here::here()
files <- dir_ls(path(project_root, "experiment_1", "full_dataset"))
demographics <- path(project_root, "clav_demographics.csv")
```

We then load in data from all of the individual data files. 
```{r message=FALSE, warning=FALSE}

#here, we create a data frame with all of the trial level data
d <- 
  files %>% 
  map_df(~read_csv(.)) %>% #read all of the CSV's into one data frame
  select(-`X16`) #get rid of this weird column that occurs in a couple of the csv's

full_participant_n <- length(unique(d$subjectID))

```

We see that there are `r full_participant_n` individual data files, although as we will see later, we are unable to recover demographic information for all of these participants.

#### Merge in demographic data

Next, we merge in demographic data.
```{r message=FALSE, warning=FALSE}
#read in demographics even though they are incomplete
demos <- 
  read_csv(demographics)

df <- 
  d %>% 
  mutate(
    is.correct = clicked == correct, 
    #add a correct column as a logical of whether 'clicked' matched 'correct'
  ) %>% 
  left_join(demos, by = "subjectID") %>% 
  mutate(
    subjectID = factor(subjectID),
  ) %>% 
  filter(!is.na(age))


demo_participant_n <- length(unique(df$subjectID))
mean_age <- mean(df$age) %>% round(2)

df <- 
  df %>% 
  mutate(age = factor(age))

```

Once we merge in the demographic information, we see that we only have full data for `r demo_participant_n` participants, meaning that we are unable to recover demographic information for `r full_participant_n - demo_participant_n` participants. Mean age of the sample is `r mean_age`. Number of participants by age and gender is shown below.

```{r}
by_age <- 
  df %>%
  group_by(subjectID) %>% 
  slice(1) %>% 
  ungroup() %>% 
  count(age, gender)

knitr::kable(by_age)

```


#### Repeated trials in some subjects

Since we notice some oddities in the data, we first look at which condition and how many trials each participant experienced.
```{r message=FALSE, warning=FALSE}
#count the number of subjects and their counterbalance conditions
subs <- 
  df %>% 
  group_by(subjectID, list_number) %>%
  count

subs
```
We find that there is some bug (?) such that occasionally identical trials are repeated. These should likely be dropped. 

### Descriptives

Next, we get a sense of the distribution of the data.
```{r message=FALSE, warning=FALSE}

summary(df)

m.rt <- exp(mean(log(df$RT), na.rm = TRUE)) 

#take log average of RT: average natural logs of RTs, then take the exponential of that

#from benny - why is he doing this log thing?

sd.rt <- exp(sd(log(df$RT),na.rm = TRUE)) #geometric standard deviation: exponentiated value of the standard deviation of the log-transformed values

qplot(RT, data = df) + 
  geom_vline(xintercept = m.rt + 2*sd.rt, col = "red", lty = 2) + #lets us view how much is spread 2 SDs above mean
  geom_vline(xintercept = m.rt - 2*sd.rt, col = "red", lty = 2) + #lets us view how much is spread 2 SDs below mean
  xlab("RT spread") +
  ylab("Frequency") +
  ggtitle("RT Histogram of CLAV Data (marking +/- 2 SDs)")
```


Here, we trim off trials with very short or long reaction times.
```{r message=FALSE, warning=FALSE}

df <- 
  df %>% 
  filter(RT > m.rt - (2 * sd.rt) & RT < m.rt + (2 * sd.rt)) #take out kids with weird RT's


```


#### Analysis: Aggregating Mean RTs and Proportion Correct

```{r message=FALSE, warning=FALSE}
summary_df <- 
  df %>% 
  group_by(subjectID, age, has_comp) %>% 
  summarize(prop_correct = mean(is.correct, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(age, has_comp) %>% 
  langcog::multi_boot_standard(col = "prop_correct")

 
summary_df %>% 
  ggplot(aes(age, fill = has_comp)) +
  geom_col(aes(y = mean), position = "dodge", width = 1) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    size = 0.3,
    width = 0.2,
    position = position_dodge(1)
  ) + #is there a better way to deal with this? 
  labs(
    y = "accuracy",
    title = "Accuracy by age and trial type",
    subtitle = "All children succeed in non-competitor trials;\n4- and 5-year-olds are more accurate than 3-year-olds on competitor trials"
  )

summary_df_no_age <- 
  df %>% 
  group_by(subjectID, has_comp) %>% 
  summarize(prop_correct = mean(is.correct, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(has_comp) %>% 
  langcog::multi_boot_standard(col = "prop_correct")

summary_df_no_age %>% 
  ggplot(aes(has_comp)) +
  geom_col(aes(y = mean), position = "dodge", width = 1) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    size = 0.3,
    width = 0.2,
    position = position_dodge(1)
  ) + #is there a better way to deal with this? 
  labs(
    y = "accuracy",
    title = "Accuracy by trial type"
  )

summary_df_no_trial_type <- 
  df %>% 
  group_by(subjectID, age) %>% 
  summarize(prop_correct = mean(is.correct, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(age) %>% 
  langcog::multi_boot_standard(col = "prop_correct")

summary_df_no_trial_type %>% 
  ggplot(aes(age)) +
  geom_col(aes(y = mean), position = "dodge", width = 1) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    size = 0.3,
    width = 0.2,
    position = position_dodge(1)
  ) + #is there a better way to deal with this? 
  labs(
    y = "accuracy",
    title = "Accuracy by age"
  )

```

Now we plot RT by trial type and age.
```{r message=FALSE, warning=FALSE}

summary_rt_df <- 
  df %>% 
  group_by(age, subjectID, has_comp) %>% 
  summarize(mean_rt = mean(RT, na.rm = TRUE)) %>% 
  ungroup() %>% 
  group_by(has_comp, age) %>% 
  langcog::multi_boot_standard(col = "mean_rt")

summary_rt_df %>% 
  ggplot(aes(x = age, y = mean, fill = has_comp)) +
  geom_col(position = "dodge", width = 1) +
  geom_errorbar(
    aes(ymin = ci_lower, ymax = ci_upper),
    size = 0.3,
    width = 0.2,
    position = position_dodge(1)
  ) +
  labs(
    title = "Reaction time by age and trial type",
    subtitle = "3-year-olds are slower to disambiguate using spoken speech than 4's and 5's",
    y = "reaction time (secs)"
  )



```

Inferential statistics:

```{r}
model <- glmer(is.correct ~ age + has_comp + age:has_comp + (has_comp | subjectID) + (1 | tar_name), data = df, family = "binomial")

summary(model)
```

The model fails to converge with age as a categorical variable.
